---
title: "Mini Literate Analysis"
author: "Leeza-marie Rodriguez"
format: html
---

## Introduction

### About This Data

The authors of this paper determined various parameters of seawater chemistry along with temperature, conductivity, and depth [data](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X) of the Arctic Ocean.

### Setup

```{r}
library(readr)
library(here)

```

### Read Data

```{r}
#| warning: false
bg_chem <- read_csv(here("data/BGchem2008data.csv"))

bg_chem <- read_csv("../data/BGchem2008data.csv")

```

Exploring Our Data

```{r}
#| eval: false
colnames(bg_chem)


str(bg_chem)

summary(bg_chem)

unique(bg_chem$Date)
```

## Analysis

### Calculate summary statistics

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
ammonium <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```

### Calculate the mean of Redfield ratio

```{r}
ratio <- (nitrate + nitrite + ammonium / phos)
```

### Plot ratio

```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

## Conclusion

The Redfield Ratio of this data is `r round(ratio)`
